### 2020-08-27 code/supplemental_v2.Rmd
Update app.R to have cancer type selection in gene and path, fix scale of last 2 plots in figs4

```
**Supplementary figure S1. Distribution of cancer type on UMAP clusters.** Each panel shows the identity of 
individual tumors by one of 23 cancer types projected onto the cluster UMAP identified 
in **Figure 2C of the main text**.

@@

# ggplot(umap_cancertype, aes(x = plot_y, y = plot_x, colour = Subtype_Selected)) + 
#   geom_point(data = umap_bg, aes(x = plot_y, y = plot_x), colour = "grey", size = 0.25) +
#   geom_point(size = 0.25) +
#   theme_classic() +
#   coord_fixed() +
#   xlab("UMAP 1") + ylab("UMAP 2") +
#   theme(axis.text.x = element_blank(),
#     axis.text.y = element_blank(),
#     axis.title.x = element_blank(),
#     axis.title.y = element_blank(),
#     axis.ticks = element_blank(),
#     axis.line = element_blank(),
#     legend.position = "none") +
#   facet_wrap(~ project_code, ncol = 5) ## facet by cancer type
```
**Supplementary Figure S2. Histological subtypes.** Each panel shows the histological subtypes of one of the 23 
cancers projected onto the tumors of the UMAP from **Figure 2C of the main text**.

@@

par(mfrow = c(2,3), mai = c(0.1, 0.1, 2, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[7:12]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[13:15]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
```
**Supplementary Figure S3. Distribution of mismatch repair genes on UMAP clusters.** Each panel shows one of 15 
genes involved in genome stability projected on the UMAP from **Figure 2C of the main text**.

@@

data <- startingpointlist[1,]
s4plot(x = data)
data <- startingpointlist[2,]
s4plot(x = data)
data <- startingpointlist[3,]
s4plot(x = data)
data <- startingpointlist[4,]
s4plot(x = data)
data <- startingpointlist[5,]
s4plot(x = data)
data <- startingpointlist[6,]
s4plot(x = data)
data <- startingpointlist[7,]
s4plot(x = data)
data <- startingpointlist[8,]
s4plot(x = data)
data <- startingpointlist[9,]
s4plot(x = data)
data <- startingpointlist[10,]
s4plot(x = data)
data <- startingpointlist[11,]
s4plot(x = data)
data <- startingpointlist[12,]
s4plot(x = data)
data <- startingpointlist[13,]
s4plot(x = data)
data <- startingpointlist[14,]
s4plot(x = data)
data <- startingpointlist[15,]
s4plot(x = data)
data <- startingpointlist[16,]
s4plot(x = data)
data <- startingpointlist[17,]
s4plot(x = data)
data <- startingpointlist[18,]
s4plot(x = data)
data <- startingpointlist[19,]
s4plot(x = data)
data <- startingpointlist[20,]
s4plot(x = data)
data <- startingpointlist[21,]
s4plot(x = data)
data <- startingpointlist[22,]
s4plot(x = data)
data <- startingpointlist[23,]
s4plot(x = data)
data <- startingpointlist[24,]
s4plot(x = data)
data <- startingpointlist[25,]
s4plot(x = data)

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[376:377]) { 
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}
```
**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2C of the main text**.

@@

```{r supplemental-s5-plot, fig.width=9}
# umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv")
# 
# umap_cancertype$clust_knn <- as.character(umap_cancertype$clust_knn)
# 
# clusters178 <- umap_cancertype %>% 
#        mutate(color = case_when(clust_knn %in% c("1", "7", "8") ~ clust_knn,
#                                   TRUE ~ "Other"))
# 
# I_outlines <- c("black", "black", "black", "#00000000")
# names(I_outlines) <- c("1", "7", "8", "Other")
# 
# I_colors <- c("gold2", "gray51", "springgreen4", "#00000011")
# names(I_colors) <- c("1", "7", "8", "Other")
# 
# load("../data/s5_data.Rda")
# genes <- as.character(unique(s5_data$variable))
# 
# par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
# par(mar = c(2,2,0,0))
# plot.new()
# with(clusters178, 
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = I_colors[color], col = I_outlines[color],
#                                 pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# legend("top", inset = c(0.05), legend=names(I_colors[c(1:3)]), col=I_colors[c(1:3)], pch = 20, bty = "n", horiz = T, cex = 1.2)
# title(xlab="UMAP 2", line=0)
# plot.new()

load("../data/s5_data.Rda")
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

@@

```
**Supplementary Figure S5. Distribution of oncogenes/tumor suppressors on UMAP clusters.** Each panel shows shows one of the twelve oncogenes/tumor suppressors projected onto the tumors of the UMAP from **Figure 2C of the main text**.

>>>>

```
**Supplementary figure S1. Distribution of cancer type on UMAP clusters.** Each panel shows the identity of 
individual tumors by one of 23 cancer types projected onto the cluster UMAP identified 
in **Figure 2 of the main text**.

@@

# ggplot(umap_cancertype, aes(x = plot_y, y = plot_x, colour = Subtype_Selected)) + 
#   geom_point(data = umap_bg, aes(x = plot_y, y = plot_x), colour = "grey", size = 0.25) +
#   geom_point(size = 0.25) +
#   theme_classic() +
#   coord_fixed() +
#   xlab("UMAP 1") + ylab("UMAP 2") +
#   theme(axis.text.x = element_blank(),
#     axis.text.y = element_blank(),
#     axis.title.x = element_blank(),
#     axis.title.y = element_blank(),
#     axis.ticks = element_blank(),
#     axis.line = element_blank(),
#     legend.position = "none") +
#   facet_wrap(~ project_code, ncol = 5) ## facet by cancer type
```
**Supplementary Figure S2. Histological subtypes.** Each panel shows the histological subtypes of one of the 23 
cancers projected onto the tumors of the UMAP from **Figure 2 of the main text**.

@@

par(mfrow = c(2,3), mai = c(0.1, 0.1, 2, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[7:12]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[13:15]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
```
**Supplementary Figure S3. Distribution of mismatch repair genes on UMAP clusters.** Each panel shows one of 15 
genes involved in genome stability projected on the UMAP from **Figure 2 of the main text**.

@@

data <- startingpointlist[1,]
s4plot(x = data)
data <- startingpointlist[2,]
s4plot(x = data)
data <- startingpointlist[3,]
s4plot(x = data)
data <- startingpointlist[4,]
s4plot(x = data)
data <- startingpointlist[5,]
s4plot(x = data)
data <- startingpointlist[6,]
s4plot(x = data)
data <- startingpointlist[7,]
s4plot(x = data)
data <- startingpointlist[8,]
s4plot(x = data)
data <- startingpointlist[9,]
s4plot(x = data)
data <- startingpointlist[10,]
s4plot(x = data)
data <- startingpointlist[11,]
s4plot(x = data)
data <- startingpointlist[12,]
s4plot(x = data)
data <- startingpointlist[13,]
s4plot(x = data)
data <- startingpointlist[14,]
s4plot(x = data)
data <- startingpointlist[15,]
s4plot(x = data)
data <- startingpointlist[16,]
s4plot(x = data)
data <- startingpointlist[17,]
s4plot(x = data)
data <- startingpointlist[18,]
s4plot(x = data)
data <- startingpointlist[19,]
s4plot(x = data)
data <- startingpointlist[20,]
s4plot(x = data)
data <- startingpointlist[21,]
s4plot(x = data)
data <- startingpointlist[22,]
s4plot(x = data)
data <- startingpointlist[23,]
s4plot(x = data)
data <- startingpointlist[24,]
s4plot(x = data)
data <- startingpointlist[25,]
s4plot(x = data)
```

```{r, fig.width = 6.7}
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0), cex.lab = 0.8)
for(n in pathorder[376:377]) { 
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=-1.7)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.11), legend=path1, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.16), legend=path2, bty = "n", horiz = T, cex = 0.8)
}
```

**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2 of the main text**.

@@

```{r supplemental-s5-plot, fig.width=9}
load("../data/s5_data.Rda")
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

@@

```
**Supplementary Figure S5. Distribution of oncogenes/tumor suppressors on UMAP clusters.** Each panel shows shows one of the twelve oncogenes/tumor suppressors projected onto the tumors of the UMAP from **Figure 2 of the main text**.
